---
title: "Initial Korean covid 19"
author: "Giselle Gonzalez"
date: "summer 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

# Introduction



### Data is from the Patient_Info dataset from South   Korea 1 Jan to 1 May 2020  
### This data was retrieved from a downloadable zip file from kaggle  
 
Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so womit these.  
state of a patient must be died or alive - we adjust that as well     
  
# Your Question - Is dying independent of sex?  


# Methods  

result is a categorical variable, sex is categorical  
result depends on sex, we use helper file  
cat~cat  

# Results  







## Graphical results - use ggplot commands to do this    
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```


This barchart shows that there are more deaths of male  
than female  

## Numerical results  

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
 This shows that Males had 3 percent and females had 1.4 percent that died   
 more than twice percent of male died then female.




## Inferential results   




## Conclusions

