---
title: "Initial Korean covid 19"
author: "Giselle Gonzalez"
date: "summer 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

# Introduction
 We will be determining if there is a difference between the sex death during covid


### Data is from the Patient_Info dataset from South   Korea 1 Jan to 1 May 2020  
### This data was retrieved from a downloadable zip file from kaggle  
 
Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so womit these.  
state of a patient must be died or alive - we adjust that as well     
  
# Your Question - Is dying independent of sex?  

$H_0$ : dying is dependent of sex during covid There is a difference depending on sex 

$H_A$ : there is no difference between the sex during covid in korea 


# Methods  

result is a categorical variable, sex is categorical  
result depends on sex, we use helper file  
cat~cat  

# Results  







## Graphical results - use ggplot commands to do this    
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```


This barchart shows that there are more deaths of male  
than female  

## Numerical results  

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
 This shows that Males had 3 percent and females had 1.4 percent that died   
 more than twice percent of male died then female.




## Inferential results   


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared test is 0.001 so we reject the null hypothesis. 
That gender and death by covid are dependent  


 The fisher exact test shows that men are 2.3 times more likley to die than women in korea  
 
 
## Conclusions

In conclusion The question was is dying independent of sex?   
We found that there was a significant difference in the graph of covid deaths of female and male because  
male are 2.3 more likely to have more deaths than females in korea. It shows that they are not related.  
